---
title: "Wrangling in R Part 1"
subtitle: "Working with Flat Files"
format: html
execute: 
  echo: true
  message: false
  warning: false
---

For this code to run, you need to have it in an .RProj file with your messy data CSV file placed in a subfolder called "data." Alternatively you can ammend the file path to exclude the `data/` part of the path and just keep the CSV file in the same folder as the .RProj file.

Run the code chunks from top to bottom. Click "Render" to render the document as an HGML file.

## Step 1: Read the WB .csv file

```{r}
library(readr)
library(dplyr)

wb_data_messy <- read_csv("data/messy_wb_data.csv") # ammend based on your file path

glimpse(wb_data_messy)
```

## Step 2: Reshape the data

```{r}
library(tidyr)
wb_data <- wb_data_messy |> 
  pivot_longer(
    cols = `1974 [YR1974]`: `2021 [YR2021]`, 
    names_to = "year", 
    values_to = "flfp"
  ) 

glimpse(wb_data)
```

Step 3: Fix year and change variables to numeric

```{r}
wb_data <- wb_data |> 
  mutate(year = substring(year, 1, 4)) |> 
  mutate_at(c("year", "flfp"), as.numeric)

glimpse(wb_data)
```

Step 4: Clean the variable names

```{r}
library(janitor)

wb_data_clean <- wb_data |>  
  clean_names() 

glimpse(wb_data)
```



